Waggle dance distributions quantify collective foraging in honey bee colonies: Supplementary material

### 2022-04-05 analysis/manuscript/paper.Rmd
updated SI and MS in text with AIC data.

In a hypothetical colony that relied only on individual search to find foraging sites, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase. By setting $p = 1$, we obtain a model based on the sole use of individual search (scouts only) and by allowing the proportion of scout trips, $p$, to take on any value between 0 and 1, we can derive the extent to which foraging occurs collectively. (Fig 2). By fitting these two models to distribution of waggle run durations decoded from real honeybee colonies we are thus able, using model selection [@Burnham2002], to infer if and, by estimating the parameter $p$, to what extent honeybee colonies use waggle dance information when foraging.

@@

For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig. 3a). In all but one of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, whereas for 9 of these sites, the distribution predicted by the scout model was significantly different to the observed data (Fig. 3b). The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (STU) and individual (ZSL) models fit best. Note the closeness of the fit to the data, illustrating the overall quality of the model description.

[Bit on AIC, link to SI showing full tables and use of delta AIC]
In all three sites where the individual model had a lower AIC than the collective model, the maximum likelihood estimate (MLE) was higher for the collective model. This indicates that the individual models do not explain the data any better than the collective model, and that the added points to the MLE comes from the increased complexity of the collective model. This is further demonstrated by the finding that the $p$ values estimated from the collective model do not settle close to 1, which would be expected if the individual model had a higher MLE, and thus suggests the collective model over fits in these sites (See supplementary material).

>>>>

In a hypothetical colony that relied only on individual search to find foraging sites, all trips are "scout" trips, but as recruitment becomes more important, the proportion of “recruit” trips will increase. By setting $p = 1$, we obtain a model based on the sole use of individual search (scouts only) and by allowing the proportion of scout trips, $p$, to take on any value between 0 and 1, we can derive the extent to which foraging occurs collectively. (Fig 2). In fitting these two models to distribution of waggle run durations decoded from real honeybee colonies we are thus able, using model selection [@Burnham2002], to infer if and, by estimating the parameter $p$, to what extent honeybee colonies use waggle dance information when foraging.

@@

For 17 out of 20 study hives, a model of collective foraging provided a better explanation of the data than one of individual foraging (Fig. 3a). In all but one of these cases, the collective model had a good fit (using a Kolmogorov-Smirnov statistic, see methods) to the empirical waggle run durations, whereas for 9 of these sites, the distribution predicted by the scout model was significantly different to the observed data (Fig. 3b). The sites STU and ZSL shown in Figs 3c-d are representative examples showing the model fits where the collective (ZSL) and individual (STU) models fit best. Note the closeness of the fit to the data, illustrating the overall quality of the model description.

In all twenty sites the collective model provided a higher maximum likelihood estimate (MLE) than the individual model, however, this is anticipated as the individual model is a subset of the collective model. In 17 of the sites the collective model accounted for over 95% of Akaike weights compared to the individual models fit (supplementary table 1). In 3 of the sites the individual model explained a higher proportion of Akaike weight than the collective model (supplementary table 1) and so honeybee foraging behaviour in these colonies is better explained by the individual rather than collective foraging model.

### 2022-04-05 analysis/manuscript/supplementary_material.Rmd
updated SI and MS in text with AIC data.

library(dplyr)
library(kableExtra)
library(flextable)
library(tibble)
library(purrr)

# load fitting results / stats
load("results/site_fit_results.Rdata")
```

@@

Fig 1. shows the models fit to all 20 sites.

```{r, all-model-fits}

plots <- map(all_sites, 2)
plots <- plots %>%
  map(
    ~ {.x + theme(
      axis.title.x = element_blank(),
      axis.title.y = element_blank()
    )}
  )
print(plots$BEL)

plot_grid(plotlist = plots, nrow = 5)

```

@@

# format table
df2 %>%
  arrange(site) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  flextable() %>%
  set_caption(
    "Fit results for the fitting of the waggle dance model to each site."
  ) %>%
  set_table_properties(layout = "autofit")
```

>>>>

library(dplyr)
library(kableExtra)
library(flextable)
library(tibble)
library(purrr)
library(gridExtra)
library(grid)

# load fitting results / stats
load("../results/site_fit_results.Rdata")
```

@@

```{r, all-model-fits, fig.width=10,fig.height=11, fig.cap="Supplementary Fig 1. log Compliment cumulative frequency distribution of waggle dance durations and fits of the collective (red line) and individual (blue line) models for all 20 sites."}
plots <- map(all_sites, 2)
plots <- plots %>%
  map(
    ~ {
      .x + theme(
        axis.title.x = element_blank(),
        axis.title.y = element_blank()
      )
    }
  )

allsites_plot <- plot_grid(
  plotlist = plots, nrow = 5,
  hjust = -0.1, label_x = 0.7,
  labels = names(plots)
)
y.grob <- textGrob(
  "cumulative frequency",
  gp = gpar(fontsize = 18),
  rot = 90
)
x.grob <- textGrob(
  "waggle run duration (s)",
  gp = gpar(fontsize = 18)
)

# add to plot
grid.arrange(arrangeGrob(allsites_plot, left = y.grob, bottom = x.grob))
```

@@

# set up proper names
# format table
df2 %>%
  arrange(site) %>%
  mutate(across(where(is.numeric), round, 3)) %>%
  rename(
    MLE = loglikelihood,
    "ΔAIC" = delta_AIC,
    D = ks_statistic,
    p_value = ks_pvalue
  ) %>%
  select(-k) %>%
  flextable() %>%
  set_caption(
    "Fit results for the fitting of the waggle dance model to each site."
  ) %>%
  set_table_properties(layout = "autofit") %>%
  fit_to_width(7.5)
```
